---
title: "Methodology: GVC Decomposition with adbmrio"
author: "Lila Ballav Bhusal"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Methodology: GVC Decomposition with adbmrio}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `adbmrio` package provides a streamlined toolkit for analyzing Global Value Chains (GVCs) using the Asian Development Bank Multi-Regional Input-Output (ADB MRIO) tables. This vignette explains the mathematical framework behind the two primary decomposition methods implemented in the package: the **National 5-Part Decomposition** (Zhang et al., 2017) and the **Bilateral WWZ Decomposition** (Wang et al., 2013).

We demonstrate these methods using the built-in `toy_mrio` dataset (4 countries, 3 sectors).

## 1. Fundamental Framework

Consider a world with $G$ countries and $N$ sectors. The global input-output system is defined by the standard Leontief equation:

$$
X = (I - A)^{-1} Y = B Y
$$

Where:

* $X$: $GN \times 1$ vector of Gross Output.
* $A$: $GN \times GN$ matrix of technical coefficients (Direct requirements).
* $Y$: $GN \times G$ matrix of Final Demand (aggregated by destination country).
* $B$: The Global Leontief Inverse ($B = (I-A)^{-1}$).

### Local vs. Global Leontief

A critical distinction in GVC analysis is between the **Global Leontief Inverse** ($B$) and the **Domestic Leontief Inverse** ($L$).

The package calculates $L$ by isolating the block-diagonal elements of $A$ (domestic transactions only, denoted as $A^D$):

$$
L = (I - A^D)^{-1}
$$

While $B$ captures feedback loops across the entire world, $L$ captures how domestic industries ripple through *only* the local economy.

---

## 2. National 5-Part Decomposition (Macro)

The function `decompose_national_5part()` breaks down a country's gross output (or Value Added/Emissions) into five distinct paths of absorption. This follows the logic proposed by Zhang et al. (2017), focusing on where production is ultimately consumed.

For a source country $s$, the total output $X_s$ is decomposed as:

$$
X_s = \underbrace{L_{ss} Y_{ss}}_{\text{Dom\_Fin}} + \underbrace{L_{ss} A^D_{ss} X_s^*}_{\text{Dom\_Int}} + \underbrace{L_{ss} \sum_{r \neq s} Y_{sr}}_{\text{Exp\_Fin}} + \underbrace{L_{ss} \sum_{r \neq s} A_{sr} L_{rr} Y_{rr}}_{\text{Exp\_Int}} + \underbrace{\text{Residual}}_{\text{Exp\_GVC}}
$$

### The Components

1. **Dom_Fin:** Output produced and consumed domestically as final goods.
2. **Dom_Int:** Output used as intermediates to produce other domestic goods consumed domestically.
3. **Exp_Fin:** Exports consumed immediately by the importer as final goods.
4. **Exp_Int:** Intermediate exports used by the importer to produce goods for their own consumption.
5. **Exp_GVC:** Intermediate exports that enter complex trade networks (re-exported to third countries or returned to the source).

### Example Usage

```{r setup}
library(adbmrio)
library(data.table)

# Load the 4-country, 3-sector toy model
data(toy_mrio)
```

```{r national_example}
# Run the decomposition for 2023
macro_results <- decompose_national_5part(toy_mrio, 2023)

# View results for China's Manufacturing Sector
china_mfg <- macro_results[country == "China" & sector == "Manufacturing"]

# Display the 5 components of Value Added
print(china_mfg[, .(VA_Dom_Fin, VA_Dom_Int, VA_Exp_Fin, VA_Exp_Int, VA_Exp_GVC)])
```

---

## 3. Bilateral WWZ Decomposition (Micro)

While the national decomposition looks at a country's total output, the `compute_bilateral_wwz()` function zooms in on the trade relationship between a specific **Source ($s$)** and **Reporting ($r$)** country.

It isolates the **Domestic Value Added (DVA)** embodied in gross exports. According to the Wang-Wei-Zhu (WWZ) framework, bilateral exports can be traced into absorption paths:

### Mathematical Formulation used in Package

Let $V_s$ be the direct value-added coefficient vector for country $s$. The package calculates:

1. **DVA in Final Goods Exports ($T_f$):**
   Value added crossing the border as a final good.
   
   $$
   DVA\_Fin = V_s L_{ss} Y_{sr}
   $$

2. **DVA in Intermediate Exports absorbed by Importer ($T_i$):**
   Value added crossing as an intermediate, processed by $r$, and consumed in $r$.
   
   $$
   DVA\_Int = V_s L_{ss} A_{sr} L_{rr} Y_{rr}
   $$

3. **DVA in GVC Trade (Re-exports) ($T_{g1}$):**
   Value added that is re-exported by $r$ to third countries.
   
   $$
   DVA\_GVC = V_s L_{ss} A_{sr} \sum_{t \neq s, r} B_{rt} Y_{t}
   $$

### Example Usage

We analyze the value-added trade flow from **China (Index 1)** to **India (Index 2)**.

```{r bilateral_example}
# 1 = China, 2 = India
bilateral_res <- compute_bilateral_wwz(toy_mrio, 2023, s_idx = 1, r_idx = 2)

print(bilateral_res)
```

### Interpretation

* **DVA_Fin:** How much Chinese value added did India consume directly as finished Chinese goods?
* **DVA_Int:** How much Chinese value added did India import as parts, turn into Indian goods, and consume?
* **DVA_GVC1:** How much Chinese value added did India import, process, and then re-export to the rest of the world?

---

## 4. Understanding the Results

### National Decomposition Output

The `decompose_national_5part()` function returns a data.table with the following columns:

* **country:** Source country name
* **sector:** Industry sector name
* **GO_*:** Gross Output components (Dom_Fin, Dom_Int, Exp_Fin, Exp_Int, Exp_GVC)
* **VA_*:** Value Added components (same breakdown)
* **CO2_*:** CO2 Emissions components (same breakdown)

This allows you to answer questions like:

* What percentage of China's manufacturing output is exported?
* How much of India's value added comes from domestic consumption vs. exports?
* Which sectors are most embedded in global value chains?

### Bilateral Decomposition Output

The `compute_bilateral_wwz()` function returns a data.table with sector-level bilateral trade flows decomposed into:

* **DVA_Fin:** Direct value added in final goods
* **DVA_Int:** Value added in intermediates absorbed by partner
* **DVA_GVC1:** Value added re-exported to third countries
* **DVA_GVC2:** Value added that returns to source country
* **FVA:** Foreign value added in exports

---

## 5. Practical Applications

### Application 1: Carbon Footprint Analysis

```{r carbon_example, eval=FALSE}
# Analyze where China's manufacturing emissions end up
china_emissions <- macro_results[country == "China" & sector == "Manufacturing"]

# Calculate export-embodied emissions
export_emissions <- china_emissions$CO2_Exp_Fin + 
                    china_emissions$CO2_Exp_Int + 
                    china_emissions$CO2_Exp_GVC

cat("China Manufacturing Export Emissions:", round(export_emissions, 2), "units\n")
```

### Application 2: Trade Policy Analysis

```{r trade_policy, eval=FALSE}
# Compare bilateral trade relationships
china_india <- compute_bilateral_wwz(toy_mrio, 2023, s_idx = 1, r_idx = 2)
china_usa <- compute_bilateral_wwz(toy_mrio, 2023, s_idx = 1, r_idx = 3)

# Which trade partner has more complex value chains?
china_india_gvc_share <- sum(china_india$DVA_GVC1) / sum(china_india$DVA_Fin + china_india$DVA_Int)
china_usa_gvc_share <- sum(china_usa$DVA_GVC1) / sum(china_usa$DVA_Fin + china_usa$DVA_Int)

cat("China-India GVC Share:", round(china_india_gvc_share * 100, 1), "%\n")
cat("China-USA GVC Share:", round(china_usa_gvc_share * 100, 1), "%\n")
```

---

## 6. Data Requirements

The package expects MRIO data in a specific format. The input should be a list containing:

* **Z:** Intermediate demand matrix (country-sector by country-sector)
* **Y:** Final demand matrix (country-sector by country)
* **VA:** Value added vector (country-sector)
* **GO:** Gross output vector (country-sector)
* **CO2:** CO2 emissions vector (country-sector, optional)
* **countries:** Vector of country names
* **sectors:** Vector of sector names
* **years:** Vector of years available

The built-in `toy_mrio` dataset demonstrates this structure.

---

## 7. Computational Notes

### Performance

For large MRIO tables (e.g., 63 countries × 35 sectors = 2,205 dimensions):

* National decomposition: ~5-10 seconds per year
* Bilateral decomposition: ~1-2 seconds per country-pair

### Memory Requirements

The package uses sparse matrix operations where possible, but large inversions $(I-A)^{-1}$ require substantial RAM:

* Small tables (< 500 dimensions): < 1 GB
* Medium tables (500-1500 dimensions): 2-8 GB
* Large tables (1500-3000 dimensions): 16+ GB

---

## References

1. **Zhang, Z., Zhu, K., & Hewings, G. J. (2017).** A multi-regional input–output analysis of the pollution haven hypothesis from the perspective of global production fragmentation. *Energy Economics*, 64, 13-23.

2. **Wang, Z., Wei, S. J., & Zhu, K. (2013).** Quantifying international production sharing at the bilateral and sector levels. *NBER Working Paper No. 19677*.

3. **Asian Development Bank (2024).** Multi-Regional Input-Output Database. Available at: https://mrio.adbx.online/

---

## Session Info

```{r session_info}
sessionInfo()
```
